---
title: "analysis"
output: html_document
date: "2023-09-21"
---

Calcs
TA changes in estuaries are related to changes in calcium ion, nitrate, ammonium, and sulfate: DTA = 2DCa2+ + DNH4+ – DNO3– – 2DSO42–
You find an oxygen increase of about 20 umol/l/h and a TA decrease of maybe 8 umol/l/h. I cannot quickly find typical O2:C ratios for SAV, so let's use a typical value of 1.5 for marine plankton. That means the DIC decline due to photosynthesis would be 13 umol/l/h. The associated DIC decline from calcification would be 8/2 = 4 umol/l/h. So I would expect a total DIC decline of 17 umol/l/h, which is about a third of what's observed.

Photosynthesis consumes ta
Calcite precipitation/calcification consumes ta
Aerobic oxidation of reduced NH4, Mn2+, Fe2+, sulfide consume ta

Aerobic respiration produces ta
Anaerobic respiration produces ta (denitrification, DNRA, sulfate reduction)
Calcite dissolution produces ta
```{r}
library(tidyverse)

load(here("Data/flux_data.RData"))
```


```{r}
compare <- flux_data %>% 
  mutate(no23_ta_mmol_m2_h = no23_rate_umol_m2_h / 1000,
         nh3_ta_mmol_m2_h = nh3_rate_umol_m2_h / 1000, 
         po4_ta_mmol_m2_h = po4_rate_umol_m2_h / 1000,
         nuts_ta_mmol_m2_h = nh3_ta_mmol_m2_h -no23_ta_mmol_m2_h - po4_ta_mmol_m2_h) %>%
  drop_na(nuts_ta_mmol_m2_h) %>%
  select(date, site, id, chamber_id, treatment, nuts_ta_mmol_m2_h, ta_rate_mmol_m2_h) %>%
  rename(`Flux from NO23, NH4, and PO4` = nuts_ta_mmol_m2_h,
         `Measured flux` = ta_rate_mmol_m2_h) %>%
  pivot_longer(6:7)
```

```{r}
ggplot(compare, aes(x = chamber_id, y = value, fill = name)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~site) +
  labs(x = "Chamber ID", y = expression("TA flux (mmol m"^-2~"h"^-1*")")) +
  theme_light()
```

