---
title: "CHALK SAV data"
date: "5/25/2023"
output: html_document
---

```{r, include = FALSE}
library(tidyverse)
library(lubridate)
library(hms)
library(here)
library(survival)
library(patchwork)

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

```{r}
#Read in DIC and TA data
ids <- read_csv(here::here("Data/sample-list.csv"))

dic <- read_csv(here::here("Data/dic_2023.csv")) %>%
  left_join(ids) %>%
  mutate(site = paste(river, site, sep = "_"),
         id = paste(site, chamber_id, sep = "_"))


ta <- read_csv(here::here("Data/ta_2023.csv")) %>%
  select(-sample) %>%
  left_join(ids) %>%
  mutate(site = paste(river, site, sep = "_"),
         id = paste(site, chamber_id, sep = "_"))
```

```{r}
#Read in oxygen logger data. Times are in EST.
do_files <- list.files(here::here("Data/loggers"), pattern = "DO")
do_names <-substr(do_files, 12, 22)
col_names = c("timestamp", "datetime_utc", "datetime_est", "battery", "temp_c", "do_mg_l", "do_pct_sat", "q")

do <- paste(here::here("Data/loggers/"), do_files, sep = "") %>%
  set_names(do_names) %>%
  map_dfr(read_csv, skip = 9, col_names = col_names, .id = "id") %>%
  select(id, datetime_est, temp_c, do_mg_l, do_pct_sat) %>%
  mutate(date = date(datetime_est),
         id = str_replace(id, pattern = "_DO", replacement = ""),
         site = substr(id, 1, 6),
         chamber_id = substr(id, 8, 8),
         datetime = round_date(datetime_est, "5 mins")) %>%
  select(-datetime_est)
```

```{r}
#Read in pH logger data and convert to EST.
ph_files <- list.files(here::here("Data/loggers"), pattern = "PH")
ph_names <-substr(ph_files, 12, 22)

ph <- paste(here::here("Data/loggers/"), ph_files, sep = "") %>%
  set_names(ph_names) %>%
  map_dfr(read_csv, .id = "id") %>%
  mutate(datetime = mdy_hm(datetime),
         date = date(datetime),
         id = str_replace(id, pattern = "_PH", replacement = ""),
         site = substr(id, 1, 6),
         chamber_id = substr(id, 8, 8),
         datetime = round_date(datetime, "5 mins")) %>%
  select(id, ph, datetime, date, site, chamber_id) %>%
  filter(id != "YOR_BR_A")
```


```{r}
#Read in HOBO logger data and convert to EST.
depth_files <- list.files(here::here("Data/loggers"), pattern = "DEPTH")
depth_names <-substr(depth_files, 12, 25)

depth <- paste(here::here("Data/loggers/"), depth_files, sep = "") %>%
  set_names(depth_names) %>%
  map_dfr(read_csv, .id = "id") %>%
  mutate(datetime = mdy_hm(datetime),
         date = date(datetime),
         id = str_replace(id, pattern = "_DEPTH", replacement = ""),
         site = substr(id, 1, 6),
         chamber_id = "Z",
         treatment = "AMBIENT",
         datetime = round_date(datetime, "5 mins"),
         depth_m = depth_m + 0.2) %>%
  select(id, depth_m, datetime, date, site, chamber_id)

depth_join <- depth %>%
  select(datetime, date, site, depth_m)
```


```{r}
#Read in PAR logger data. Times are in EST.
par_files <- list.files(here::here("Data/loggers"), pattern = "PAR")
par_names <-substr(par_files, 12, 23)
par_col_names = c("timestamp", "datetime_utc", "datetime_est", "battery", "temp_c", "par_umol_m2_s", "acc_x", "acc_y", "acc_z")

par <- paste(here::here("Data/loggers/"), par_files, sep = "") %>%
  set_names(par_names) %>%
  map_dfr(read_csv, skip = 7, col_names = par_col_names, .id = "id") %>%
  select(id, datetime_est, par_umol_m2_s) %>%
  mutate(date = date(datetime_est),
         id = str_replace(id, pattern = "_PAR", replacement = ""),
         site = substr(id, 1, 6),
         chamber_id = "Z",
         datetime = round_date(datetime_est, "5 mins")) %>%
  select(-datetime_est)
```


```{r}
#Read in chamber info and SAV data. Convert times to EST.
times <- read_csv(here::here("Data/sample_times.csv")) %>%
  mutate(id = paste(river, site, chamber_id, sep = "_"),
         site = paste(river, site, sep = "_")) %>%
  drop_na(date) %>%
  select(date, T1, T2, id, site, chamber_id, treatment) %>%
  mutate(T1 = ymd_hms(paste(date, T1, sep = " ")),
         T2 = ymd_hms(paste(date, T2, sep = " ")),
         T1 = T1 + minutes(15),
         T2 = T2 - minutes(5))

biomass <- read_csv(here::here("Data/biomass.csv")) %>%
  mutate(id = paste(river, site, chamber_id, sep = "_"),
         site = paste(river, site, sep = "_")) %>%
  drop_na(date) %>%
  filter(ag_bg != "animal")

sav_weight <- read_csv(here::here("Data/biomass.csv")) %>%
  mutate(id = paste(river, site, chamber_id, sep = "_"),
         site = paste(river, site, sep = "_")) %>%
  drop_na(date) %>%
  group_by(date, id, site, chamber_id, ag_bg) %>%
  summarize(sav_wt = sum(dry_mass_g)) %>%
  filter(ag_bg == "ag") %>%
  select(-ag_bg)

vols <- read_csv(here::here("Data/chamber_volume_calcs.csv")) %>%
  mutate(id = paste(river, site, chamber_id, sep = "_")) %>%
  drop_na(date) %>%
  select(date, id, vol_l)
```

```{r}
#get pH at T1 and T2 for calcium samples

ph %>% 
  select(date, datetime, ph, id) %>%
  left_join(times) %>%
  filter(date > ymd("2023-07-31")) %>%
  group_by(date, id, site, chamber_id, treatment) %>% 
  filter(abs(difftime(datetime,T2)) == min(abs(difftime(datetime,T2)))) %>%
  filter(chamber_id %in% c("A", "B","D", "E", "G", "J")) %>%
  print(n = 40)
```

```{r}
#Join logger data for time series plots.
logger_join <- do %>%
  full_join(ph) %>%
  full_join(depth) %>%
  full_join(par) %>%
  full_join(times, multiple = "all") %>%
  group_by(date, site, treatment, chamber_id) %>%
  filter(datetime >= T1 & datetime <= T2)
```

**Abbreviations**

Sites:

- POT_BR: Potomac brackish (Piney Point) 
- POT_FW: Potomac freshwater (Chickamuxen Creek)
- YOR_BR: York brackish (Guinea marsh)
- YOR_FW: York freshwater (Pamunkey near Chamberlayne Point) 

Treatments:
 
- AMBIENT: Ambient
- SAV-D: SAV dark  
- SAV-L: SAV light 
- SED-D: sediment + water dark
- SED-L: sediment + water light 
- W-D: water dark 
- W-L: water light 
 
```{r}
#Oxygen time series.
ggplot(logger_join, aes(x = datetime, y = do_mg_l, color = chamber_id)) +
  geom_line() +
  facet_grid(treatment~site, scales = "free_x") +
  labs(x = "", y = expression("DO (mg l"^-1*")"),
       title = "Oxygen time series") +
  theme_light() +
  theme(legend.position = "none")
```

Oxygen time series are generally as expected, with increases in light chambers - especially those containing SAV - due to primary production and decreases in dark chambers due to respiration. A notable exception are the YOR-BR water treatments. Temperature and water depth also changed, which could have affected oxygen saturation. We also noted that it appeared that the two chamber bags may have been leaking.

```{r}
ggplot(logger_join, aes(x = datetime, y = ph, color = chamber_id)) +
  geom_line() +
  facet_grid(treatment~site, scales = "free_x") +
  labs(x = "", y = "pH",
       title = "pH time series") +
  theme_light() +
  theme(legend.position = "none")
```

YOR_FW is missing pH data due to human logger programming error!

```{r}
ggplot(logger_join, aes(x = datetime, y = temp_c, color = chamber_id)) +
  geom_line() +
  facet_grid(treatment~site, scales = "free_x") +
  labs(x = "", y = expression("Temp. ("*degree*"C)"),
       title = "Temperature time series") +
  theme_light() +
  theme(legend.position = "none")

ggplot(logger_join, aes(x = datetime, y = depth_m)) +
  geom_path() +
  facet_wrap(~site, scales = "free_x") +
  labs(x = "", y = "Depth (m)",
       title = "Water level") +
  theme_light()

ggplot(logger_join, aes(x = datetime, y = par_umol_m2_s)) +
  geom_path() +
  facet_wrap(~site, scales = "free_x") +
  labs(x = "", y = expression("PAR ("*mu*"mol m"^-2~"s"^-1*")"),
       title = "PAR at the bottom") +
  theme_light()
```

POT_BR and YOR_FW were both deeper and more turbid than YOR_BR.

```{r}
#plot biomass
biomass %>%
  mutate(ag_bg = case_when(ag_bg == "ag" ~ "above ground",
                           ag_bg == "bg" ~ "below ground"),
         species = case_when(species == "cd" ~ "Ceratophyllem demersum",
                             species == "hv" ~ "Hydrilla verticillata",
                             species == "lyng" ~ "Lyngbya spp.",
                             species == "ng" ~ "Najas guadeloupensis",
                             species == "red_algae" ~ "red algae",
                             species == "cladophora" ~ "green algae",
                             species == "rm" ~ "Ruppia maritima",
                             species == "zm" ~ "Zostera marina",
                             species == "unknown" ~ "mixed species roots")) %>%
ggplot(aes(x = chamber_id, y = biomass_g_m2, fill = species)) +
  geom_bar(stat = "identity") +
  facet_grid(ag_bg ~site) +
  labs(x = "Chamber ID", y = expression("SAV biomass (g m"^-2*")"),
       title = "Chanmber SAV biomass") +
  theme_light()
```


```{r}
#Calculate O2 flux rates.
o2_change <- function(do){
  delta_o <- rep(NA, length(do)) 
  for(i in 2:length(do)) {delta_o[i] = do[i] - do[i-1]}
  delta_o
}

fluxes <- do %>%
  full_join(times) %>%
  left_join(depth_join) %>%
  group_by(date, id, site, treatment, chamber_id) %>%
  filter(datetime >= T1 & datetime <= T2) %>%
  select(-c(T1, T2)) %>%
  mutate(flux_mg_l = o2_change(do_mg_l),
        delta_t_h = 0.08333,
        flux_water_mg_m2 =if_else(str_detect(treatment, "W"), 
                                  (flux_mg_l * depth_m) * 1000,
                                  NA)) %>%
  summarize(time_h = sum(delta_t_h),
            rate_mg_l_h = sum(flux_mg_l, na.rm = TRUE)/time_h,
            flux_water_mg_m2_h = sum(flux_water_mg_m2, na.rm = TRUE)/time_h) %>%
  full_join(sav_weight) %>%
  full_join(vols) %>%
  mutate(rate_mmol_l_h = rate_mg_l_h/32,
         rate_mmol_m2_h = if_else(str_detect(treatment, "W"),
                             flux_water_mg_m2_h / 32,
                             rate_mmol_l_h * vol_l / 0.018),
         rate_mmol_g_sav_h = rate_mmol_l_h * vol_l / sav_wt) %>%
  select(-c(flux_water_mg_m2_h)) %>%
  drop_na(site)
```

```{r}
#Plot fluxes (mg/l/h)
ggplot(fluxes, aes(x = chamber_id, y = rate_mg_l_h, fill = treatment)) +
  geom_bar(stat="identity") +
  facet_wrap(~site) +
  labs(x = "Chamber ID", y = expression("O"[2]~"flux (mg l"^-1~"h"^-1*")"),
       title = "Oxygen fluxes",
       subtitle = "(not normalized to biomass or chamber volume)") +
  theme_light()
```

General oxygen flux patterns are as expected, with the exception of YOR_BR W-D which indicates oxygen production in the dark. 

```{r}
#calculate change in pH
ph_change <- function(ph){
  delta_ph <- rep(NA, length(ph)) 
  for(i in 2:length(ph)) {delta_ph[i] = ph[i] - ph[i-1]}
  delta_ph
}

delta_ph <- logger_join %>%
  select(date, datetime, id, site, treatment, chamber_id, ph) %>%
  full_join(times) %>%
  mutate(T1 = ceiling_date(T1, "5 mins"),
         T2 = floor_date(T2, "5 mins")) %>%
  group_by(date, id, site, treatment, chamber_id) %>%
  filter(datetime == T1 | datetime == T2) %>%
  select(-c(T1, T2)) %>%
  mutate(ph_change = ph_change(ph)) %>%
  drop_na(ph_change)
```


```{r}
#Plot delta pH
ggplot(delta_ph, aes(x = chamber_id, y = ph_change, fill = treatment)) +
  geom_bar(stat = "identity") +
  facet_wrap(~site) +
  labs(x = "Chamber ID", y = expression(Delta~"pH"),
       title = expression("pH change (pH"[T2]~"- pH"[T1]*")"),
       subtitle = "during incubation (2-4 h)") +
  theme_light() 
```


```{r}
#Calculate DIC change
dic_flux <- dic %>%
  select(id, site, treatment, replicate, chamber_id, dic_mmol, sample) %>%
  pivot_wider(names_from = sample, values_from = dic_mmol) %>%
  mutate(dic_change_mmol_l = T2 - T1) %>%
  select(-T1, -T2) %>%
  full_join(times) %>%
  full_join(vols) %>%
  mutate(time_change = T2 - T1,
         flux_mmol_l_h = dic_change_mmol_l / as.numeric(time_change),
         flux_mmol_m2_h = flux_mmol_l_h * vol_l) %>%
  select(-T1, -T2) %>%
  filter(site %in% c("YOR_BR", "YOR_FW"))
```


```{r}
#Calculate TA change
ta_flux <- ta %>%
  select(id, site, treatment, replicate, chamber_id, ta_umol_l, sample) %>%
  pivot_wider(names_from = sample, values_from = ta_umol_l) %>%
  mutate(ta_change_umol_l = T2 - T1) %>%
  select(-T1, -T2) %>%
  full_join(times) %>%
  full_join(vols) %>%
  mutate(time_change = T2 - T1,
         flux_umol_l_h = ta_change_umol_l / as.numeric(time_change)) %>%
#         flux_umol_m2_h = flux_umol_l_h * vol_l) %>%
  select(-T1, -T2) %>%
  filter(site %in% c("YOR_BR", "YOR_FW"))
```

```{r}
#plot DIC TA and DO
dicplot <- ggplot(dic_flux, aes(x = chamber_id, y = flux_mmol_l_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  facet_wrap(~site) +
  lims(y = c(-0.068, 0.026)) +
  labs(x = "Chamber ID", y = expression(Delta~"DIC (mmol l"^-1~"h"^-1*")"),
       title = "DIC flux") +
  theme_light()

taplot <- ggplot(ta_flux, aes(x = chamber_id, y = flux_umol_l_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  facet_wrap(~site) +
  labs(x = "Chamber ID", y = expression(Delta~"TA (umol l"^-1~"h"^-1*")"),
       title = "TA flux") +
  theme_light()

doplot <- fluxes %>%
  filter(site %in% c("YOR_BR", "YOR_FW"),
         chamber_id != "Z") %>%
  ggplot(aes(x = chamber_id, y = rate_mmol_l_h, fill = treatment)) +
  geom_bar(stat="identity") +
  lims(y = c(-0.068, 0.026)) +
  facet_wrap(~site) +
  labs(x = "Chamber ID", y = expression(Delta~"O"[2]~"(mmol l"^-1~"h"^-1*")"),
       title = "Oxygen flux") +
  theme_light()

phplot <- delta_ph %>%
  filter(site %in% c("YOR_BR", "YOR_FW"),
         chamber_id != "Z") %>%
  ggplot(aes(x = chamber_id, y = ph_change, fill = treatment)) +
  geom_bar(stat = "identity") +
  facet_wrap(~site) +
  labs(x = "Chamber ID", y = expression(Delta~"pH"),
       title = expression("pH change")) +
  theme_light() 

taplot + dicplot + doplot +  phplot + plot_layout(guides = 'collect') + plot_annotation(caption = "D = dark, L = light; missing values = samples not yet analyzed or sonde programming error.\nFluxes have not been normalized to chamber volume. DIC data are provisional.")

ggsave("sav_ta_dic_do_ph.png", width = 10, height = 6, units = "in", dpi = 300)
```



Missing pH change for YOR_FW due to human error.

```{r}
#Calculate mean and SE O2 flux/m2

se <- function(x) sd(x, na.rm = TRUE) / sqrt(length(x))

means_area <- fluxes %>%
  filter(treatment != "AMBIENT") %>%
  group_by(date, site, treatment) %>%
  summarize(mmol_m2_h_mean = mean(mmol_m2_h))  %>%
  separate(treatment, into = c("trtmt", "light_dark")) %>%
  pivot_wider(names_from = light_dark, values_from = mmol_m2_h_mean) %>%
  mutate(R = D,
         NPP = L, 
         GPP = L - D) %>%
  select(-c(D, L)) %>%
  pivot_longer(4:6, values_to = "means")

ses_area <- fluxes %>%
  filter(treatment != "AMBIENT") %>%
  group_by(date, site, treatment) %>%
  summarize(mmol_m2_h_se = se(mmol_m2_h))  %>%
  separate(treatment, into = c("trtmt", "light_dark")) %>%
  pivot_wider(names_from = light_dark, values_from = mmol_m2_h_se) %>%
  mutate(R = D,
         NPP = L, 
         GPP = L - D) %>%
  select(-c(D, L)) %>%
  pivot_longer(4:6, values_to = "ses")

flux_summary_area <- means_area %>%
  left_join(ses_area)
```


```{r}
#Calculate mean and SE O2 flux per g SAV

means_sav <- fluxes %>%
  filter(treatment != "AMBIENT") %>%
  group_by(date, site, treatment) %>%
  summarize(mmol_g_sav_h_mean = mean(mmol_g_sav_h, na.rm = TRUE))  %>%
  separate(treatment, into = c("trtmt", "light_dark")) %>%
  pivot_wider(names_from = light_dark, values_from = mmol_g_sav_h_mean) %>%
  mutate(R = D,
         NPP = L, 
         GPP = L - D) %>%
  select(-c(D, L)) %>%
  pivot_longer(4:6, values_to = "means") %>%
  drop_na(means)

ses_sav <- fluxes %>%
  filter(treatment != "AMBIENT") %>%
  group_by(date, site, treatment) %>%
  summarize(mmol_g_sav_h_mean = se(mmol_g_sav_h))  %>%
  separate(treatment, into = c("trtmt", "light_dark")) %>%
  pivot_wider(names_from = light_dark, values_from = mmol_g_sav_h_mean) %>%
  mutate(R = D,
         NPP = L, 
         GPP = L - D) %>%
  select(-c(D, L)) %>%
  pivot_longer(4:6, values_to = "ses") %>%
drop_na(ses)

flux_summary_sav <- means_sav %>%
  left_join(ses_sav)
```


```{r}
#Plot mean aerial flux.
ggplot(flux_summary_area, aes(x = trtmt, y = means)) +
   geom_point() + 
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_errorbar(aes(ymin = means - ses, ymax = means + ses), width = 0.2) +
   facet_grid(site ~ name) +
   labs(x = "", y = expression("DO flux (mmol O"[2]~"m"^-2~"h"^-1*")"),
        title = expression("Mean aerial O"[2]~"flux")) +
   theme_light()
```

Mean aerial oxygen flux, which normalizes fluxes to chamber volume and extrapolates to square meters. SAV and sediment rates are not included for the brackish sites because we used bromide as a tracer to measure chamber volume. The bromide samples need to be processed by CBL NASL. We used NaCl/conductivity to measure chamber volume at the freshwater sites which could analyze immediately. Water column (W) rates are calculated based on the sum of water chamber oxygen change at each time step multiplied by incubation time and water depth at each time step. The YOR_BR W rates are questionable - we either need to correct the oxygen concentrations for changes in temperature and water depth or discard the data assuming chamber leakage.

```{r}
#Plot mean SAV-mass flux.

ggplot(flux_summary_sav, aes(x = name, y = means)) +
   geom_point() + 
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_errorbar(aes(ymin = means - ses, ymax = means + ses), width = 0.1) +
  facet_wrap(~site) +
   labs(x = "", y = expression("Oxygen"~"flux (mmol O"[2]~"g SAV"^-1~"h"^-1*")"),
        title = expression("Mean oxygen flux per g SAV dry weight")) +
   theme_light()
```

Oxygen flux per g SAV biomass (dry weight). To calculate this, we multiplied oxygen change (mg/l) by chamber water volume (l), converted to mg to mmol, and divided by SAV dry weight in each chamber (g). Again, only available right now for the freshwater site because chamber volume is used in this calculation. These rates are whole-system metabolism - we didn't subract out water and sediment rates.
