---
title: "plots"
output: html_document
date: "2023-08-23"
---

```{r, include = FALSE}
library(tidyverse)
library(khroma)
library(ggpubr)
library(ggsci)
library(here)
library(patchwork)

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

```{r}
load(here("Data/flux_data.RData"))

licor <- read_csv(here("Data/par.csv"))

animals <- read_csv(here("Data/animals_epiphytes.csv")) %>%
  mutate(site = paste(river, site, sep = "_"),
         id = paste(site, chamber_id, sep = "_")) %>%
  group_by(date, id, species_id) %>%
  summarize(dry_mass_g = sum(dry_mass_g, na.rm = TRUE),
            organic_g = sum(organic_g, na.rm = TRUE), 
            inorgainc_g = sum(inorganic_g, na.rm = TRUE))

sediment <- read_csv(here("Data/biomass.csv")) %>%
  filter(ag_bg == "sediment") %>%
  mutate(id = paste(river, site, chamber_id, sep = "_")) %>%
  group_by(date, id, species) %>%
  summarize(dry_mass = sum(dry_mass_g, na.rm = TRUE),
            dry_mass_aerial = sum(biomass_g_m2, na.rm = TRUE))

biomass <- read_csv(here("Data/biomass.csv")) %>%
  filter(ag_bg != "sediment") %>%
  mutate(id = paste(river, site, chamber_id, sep = "_")) %>%
  group_by(date, id, site, chamber_id, species, ag_bg) %>%
  summarize(dry_mass_g = sum(dry_mass_g, na.rm = TRUE),
            biomass_g_m2 = sum(biomass_g_m2, na.rm = TRUE)) %>%
  mutate(season = case_when(month(date) %in% 5:6 ~ "spring",
                            month(date) %in% 7:8 ~ "summer",
                            month(date) %in% 9:10 ~ "fall"),
         year = year(date)) %>%
  ungroup()
```

RAW DATA

Compare mg/l and mg/m2 DIC
```{r}
p1 <- ggplot(flux_data, aes(x = chamber_id, y = dic_rate_mmol_m2_h, color = treatment)) +
  geom_bar(stat = "identity") +
  labs(x = "Chamber ID", y = expression("DIC flux (mmol m"^-2~"h"^-1*")")) +
  facet_grid(season~site) +
    scale_fill_npg() +
  theme_light() +
  theme(legend.position = "none")

p2 <- ggplot(flux_data, aes(x = chamber_id, y = dic_rate_mmol_l_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  labs(x = "Chamber ID", y = expression("DIC flux (mmol l"^-2~"h"^-1*")")) +
  facet_grid(season~site) +
    scale_fill_npg() +
  theme_light() +
  theme(legend.position = "none")

p1 + p2
```

Compare mg/l and mg/m2 TA
```{r}
p1 <- ggplot(flux_data, aes(x = chamber_id, y = ta_rate_mmol_m2_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  labs(x = "Chamber ID", y = expression("TA flux (mmol m"^-2~"h"^-1*")")) +
  facet_grid(season~site) +
  scale_fill_npg() +
  theme_light() +
  theme(legend.position = "none")

p2 <- ggplot(flux_data, aes(x = chamber_id, y = ta_rate_umol_l_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  labs(x = "Chamber ID", y = expression("TA flux (umol l h"^-1*")")) +
  facet_grid(season~site) +
  scale_fill_npg() +
  theme_light() +
  theme(legend.position = "none")

p1 + p2
```

Compare mg/l and mg/m2 DO
```{r}
p1 <- flux_data %>%
  filter(chamber_id != "Z") %>%
ggplot(aes(x = chamber_id, y = o2_rate_mmol_m2_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  labs(x = "Chamber ID", y = expression("O"[2]~"flux (mmol m"^-2~"h"^-1*")")) +
  facet_grid(season~site, scales = "free_y") +
  scale_fill_npg() +
  theme_light() +
  theme(legend.position = "none")

p2 <- flux_data %>%
  filter(chamber_id != "Z") %>%
ggplot(aes(x = chamber_id, y = o2_rate_mg_l_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  labs(x = "Chamber ID", y = expression("O"[2]~"flux (mg l h"^-1*")")) +
  facet_grid(season~site, scales = "free_y") +
  scale_fill_npg() +
  theme_light() +
  theme(legend.position = "none")

p1 + p2
```


```{r}
#Plot DIC, TA, and oxygen fluxes
ggplot(flux_data, aes(x = chamber_id, y = dic_rate_mmol_m2_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  labs(x = "Chamber ID", y = expression("DIC flux (mmol m"^-2~"h"^-1*")")) +
  facet_grid(season~site) +
    scale_fill_npg() +
  theme_light()

ggplot(flux_data, aes(x = chamber_id, y = ta_rate_mmol_m2_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  labs(x = "Chamber ID", y = expression("TA flux (mmol m"^-2~"h"^-1*")")) +
  facet_grid(season~site) +
  scale_fill_npg() +
  theme_light()

flux_data %>%
  filter(chamber_id != "Z") %>%
ggplot(aes(x = chamber_id, y = o2_rate_mmol_m2_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  labs(x = "Chamber ID", y = expression("O"[2]~"flux (mmol m"^-2~"h"^-1*")")) +
  facet_grid(season~site, scales = "free_y") +
  scale_fill_npg() +
  theme_light()

flux_data %>%
  filter(chamber_id != "Z") %>%
ggplot(aes(x = chamber_id, y = ph_change, fill = treatment)) +
  geom_bar(stat = "identity") +
  labs(x = "Chamber ID", y = expression(Delta~"pH")) +
  facet_grid(season~site, scales = "free_y") +
  scale_fill_npg() +
  theme_light()

ggplot(flux_data, aes(x = chamber_id, y = no23_rate_mmol_m2_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  labs(x = "Chamber ID", y = expression("NO"[2+3]~"N flux ("*mu*"mol m"^-2~"h"^-1*")")) +
  facet_grid(season~site) +
    scale_fill_npg() +
  theme_light()

ggplot(flux_data, aes(x = chamber_id, y = nh3_rate_mmol_m2_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  labs(x = "Chamber ID", y = expression("NH"[3]~"N flux ("*mu*"mol m"^-2~"h"^-1*")")) +
  facet_grid(season~site) +
    scale_fill_npg() +
  theme_light()

ggplot(flux_data, aes(x = chamber_id, y = po4_rate_mmol_m2_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  labs(x = "Chamber ID", y = expression("PO"[4]^"3-"~"P flux ("*mu*"mol m"^-2~"h"^-1*")")) +
  facet_grid(season~site) +
    scale_fill_npg() +
  theme_light()
```


```{r}
#plot biomass
biomass_plot <- biomass %>%
  complete(date, id, species, year, season, ag_bg, site, chamber_id, 
           fill = list(dry_mass_g = 0, biomass_g_m2 = 0), explicit = TRUE) %>%
  mutate(site = substr(id, 1, 6),
         season = factor(season, levels = c("spring", "summer", "fall"))) %>%
  ungroup() %>%
  mutate(ag_bg = case_when(ag_bg == "ag" ~ "above ground",
                           ag_bg == "bg" ~ "below ground"),
         species = case_when(species == "cd" ~ "Ceratophyllem demersum",
                             species == "hv" ~ "Hydrilla verticillata",
                             species == "lw" ~ "Lyngbya spp.",
                             species == "ng" ~ "Najas guadeloupensis",
                             species == "red_algae" ~ "red algae",
                             species == "cladophora" ~ "green algae",
                             species == "rm" ~ "Ruppia maritima",
                             species == "zm" ~ "Zostera marina",
                             species == "unknown" ~ "mixed species roots",
                             species == "va" ~ "Vallisneria americana",
                             species == "ms" ~ "Myriophyllem spicatum",
                             species == "pper" ~ "Potamogeton perfoliatus",
                             species == "ppus" ~ "Potamogeton pusilus",
                             species == "hd" ~ "Heteranthera dubia",
                             species == "nm" ~ "Najas minor")) %>%
    filter(ag_bg == "above ground")

nsavspecies <- length(unique(biomass$species))

ggplot(biomass_plot, aes(x = chamber_id, y = biomass_g_m2, fill = species)) +
  geom_bar(stat = "identity") +
  facet_grid(season ~ site) +
  labs(x = "Chamber ID", y = expression("SAV above ground biomass (g m"^-2*")")) +
  theme_light() +
    scale_fill_manual(values = get_palette(pal_npg("nrc")(10), nsavspecies))
```


```{r}
#Plot epi/animal raw data
animals_plot <- animals %>%
  filter(species_id != "sav") %>%
  pivot_longer(4:6, names_to = "param", values_to = "value") %>%
  group_by(date, id, species_id, param) %>%
  summarize(value = sum(value)) %>%
  ungroup() %>%
  pivot_wider(names_from = param, values_from = value) %>%
  mutate(calcifier = case_when(species_id %in% c("clam", "crab", "epi", "insect", "snail") ~ "yes",
                               TRUE ~ "no")) %>%
  select(-dry_mass_g) %>%
  rename(tissue = organic_g, shell = inorgainc_g) %>%
  pivot_longer(4:5, names_to = "param", values_to = "value") %>%
  complete(date, id, species_id, param, calcifier, 
           fill = list(value = 0), explicit = TRUE) %>%
  ungroup() %>%
  mutate(season = case_when(month(date) %in% 5:6 ~ "spring",
                            month(date) %in% 7:8 ~ "summer",
                            month(date) %in% 9:10 ~"fall"),
         season = factor(season, levels = c("spring", "summer", "fall")),
         chamber_id = substr(id, 8, 8),
         site = substr(id, 1, 6))

animals_plot %>%
  filter(calcifier == "yes") %>%
  drop_na(species_id) %>%
  pivot_wider(names_from = param, values_from = value) %>%
  ggplot(aes(x = chamber_id, y = shell, fill = species_id)) +
    geom_bar(stat = "identity") +
    labs(x = "Chamber ID", y = "Calcifier shell (g)") +
    facet_grid(season ~ site) +
  theme_light() +
    scale_fill_manual(values = get_palette(pal_npg("nrc")(7), nsavspecies))

animals_plot %>%
  filter(calcifier == "yes") %>%
  pivot_wider(names_from = param, values_from = value) %>%
  ggplot(aes(x = chamber_id, y = tissue, fill = species_id)) +
    geom_bar(stat = "identity") +
    labs(x = "Chamber ID", y = "Calcifier tissue (g)") +
    facet_grid(season ~ site) +
  theme_light() +
    scale_fill_manual(values = get_palette(pal_npg("nrc")(7), nsavspecies))

animals_plot %>%
  filter(calcifier == "no") %>%
  pivot_wider(names_from = param, values_from = value) %>%
  ggplot(aes(x = chamber_id, y = tissue, fill = species_id)) +
    geom_bar(stat = "identity") +
    labs(x = "Chamber ID", y = "Other animal tissue (g)") +
    facet_grid(season ~ site) +
  theme_light() +
    scale_fill_manual(values = get_palette(pal_npg("nrc")(7), nsavspecies))
```


```{r}
#Join flux and calcifiers dataframes
calcifier_join <- animals %>%
  filter(species_id != "sav") %>%
  pivot_longer(4:6, names_to = "param", values_to = "value") %>%
  mutate(calcifier = case_when(species_id %in% c("clam", "crab", "epi", "insect", "snail") ~ "yes",
                               TRUE ~ "no")) %>%
  group_by(date, id, calcifier, param) %>%
  summarize(value = sum(value)) %>%
  ungroup() %>%
  pivot_wider(names_from = param, values_from = value) %>%
  filter(calcifier == "yes") %>%
  select(-calcifier, -dry_mass_g) %>%
  rename(calcifier_tissue_g = organic_g, calcifier_shell_g = inorgainc_g)

worm_join <- animals %>%
  filter(species_id != "sav") %>%
  pivot_longer(4:6, names_to = "param", values_to = "value") %>%
  mutate(calcifier = case_when(species_id %in% c("clam", "crab", "epi", "insect", "snail") ~ "yes",
                               TRUE ~ "no")) %>%
  group_by(date, id, calcifier, param) %>%
  summarize(value = sum(value)) %>%
  ungroup() %>%
  pivot_wider(names_from = param, values_from = value) %>%
  filter(calcifier == "no") %>%
  select(-calcifier, -dry_mass_g, -inorgainc_g) %>%
  rename(worm_tissue_g = organic_g)

epi <- animals %>%
  filter(species_id == "epi") %>%
  select(date, id, dry_mass_g) %>%
  rename(epi_g = dry_mass_g)

flux_data <- flux_data %>%
  left_join(calcifier_join) %>%
  left_join(worm_join) %>%
  left_join(epi) %>%
  mutate(site = substr(id, 1, 6),
    Site = case_when(site == "POT_BR" ~ "Potomac\nbrackish",
                          site == "YOR_BR" ~ "York\nbrackish",
                          site == "YOR_FW" ~ "York\nfresh",
                          site == "POT_FW" ~ "Potomac\nfresh")) %>%
  mutate(sav = treatment) %>%
  separate(sav, into = c("sav", "light_dark")) %>%
  mutate(light_dark = case_when(light_dark == "D" ~ "dark",
                                light_dark == "L" ~ "light"),
         sav = case_when(sav == "SAV" ~ "vegetated",
                         sav == "SED" ~ "bare"))
```


Boxplots: 
TA: spring dark less TA consumption in SAV; light more TA consumption excepth for York fresh. Summer dark Potomac: more TA consumption in SAV, york smaller difference and not consistent. Light more TA production in SAV, excepth for York Fresh.
DIC dark more DIC production in the dark compared to bare (higher rates of respiration); more consumtption in the light (higher rates of photosynthesis)
```{r}
flux_data %>%
  drop_na(sav) %>%
  filter(season != "fall") %>%
  mutate(ta_rate_mmol_m2_h = case_when(ta_rate_mmol_m2_h < -10 ~ NA,
                                       TRUE ~ ta_rate_mmol_m2_h)) %>%
ggplot(aes(x = Site, y = ta_rate_mmol_m2_h, fill = sav)) +
  geom_hline(yintercept = 0, color = "darkgray", lty = "dashed") +
  geom_boxplot(alpha = 0.8) +
  geom_jitter(alpha = 0.5, position = position_jitterdodge(jitter.width = 0.15)) +
  facet_grid(season~light_dark) +
  labs(x = "", y = expression("TA flux (mmol m"^-2~"h"^-1*")"),
       fill = "", title = expression("Light and dark TA flux in bare and vegetated sediment")) +
  scale_fill_npg() +
  theme_light() +
  theme(legend.position = "bottom")

flux_data %>%
  drop_na(sav) %>%
  filter(season != "fall") %>%
ggplot(aes(x = Site, y = dic_rate_mmol_m2_h, fill = sav)) +
  geom_hline(yintercept = 0, color = "darkgray", lty = "dashed") +
  geom_boxplot(alpha = 0.8) +
  geom_jitter(alpha = 0.5, position = position_jitterdodge(jitter.width = 0.15)) +
  facet_grid(season~light_dark) +
  labs(x = "", y = expression("DIC flux (mmol m"^-2~"h"^-1*")"),
       fill = "", title = expression("Light and dark DIC flux in bare and vegetated sediment")) +
  scale_fill_npg() +
  theme_light() +
  theme(legend.position = "bottom")
```

*Biomass effect on TA* 
In dark, there is a generally positive effect (more biomass, more TA production, although this is mostly driven by high rates of production in Potomac Fresh). In light, there is generally a negative relationship (more biomass, more TA consumption). However, in Potomac fresh, TA production increases with biomass. 
```{r}
flux_data %>% filter(treatment != "AMBIENT") %>%
    mutate(ta_rate_mmol_m2_h = case_when(ta_rate_mmol_m2_h < -10 ~ NA,
                                         
                                       TRUE ~ ta_rate_mmol_m2_h)) %>%
ggplot(aes(x = ag, y = ta_rate_mmol_m2_h,color = Site)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(size = 3) +
  labs(x = expression("SAV biomass (g m"^-2*")"), 
       y = expression("TA flux (mmol m"^-2~"h"^-1*")"),
       color = "",
       title = "Biomass effect on TA", shape = "") +
  facet_wrap(~light_dark) +
  stat_regline_equation(aes(label =  paste(..eq.label..)), label.y = c(16, 17, 18, 19, 16, 17, 18, 19)) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")), 
           label.x = 4, label.y = c(16, 17, 18, 19, 16, 17, 18, 19)) +
  scale_color_jama() +
  theme_light() +
  theme(legend.position="bottom", legend.box="vertical", legend.margin=margin(), 
        text = element_text(size=20)) 
```

*Oxygen effect on TA flux*
No effect in POT BR. Positive effect in PO_FW. Slightly negative effect in York BR. York FW, positive effect in dark, negative effect in light.
```{r}
flux_data %>%
  drop_na(light_dark) %>%
  filter(treatment %in% c("SAV-D", "SAV-L", "SED-D", "SED-L")) %>%
  mutate(ta_rate_mmol_m2_h = case_when(ta_rate_mmol_m2_h < -10 ~ NA,
                                       TRUE ~ ta_rate_mmol_m2_h)) %>%
#  separate(Site, into = c("river", "salinity")) %>%
ggplot(aes(x = mean_o2, y = ta_rate_mmol_m2_h, color = site, shape = sav)) +
  stat_regline_equation(aes(label =  paste(..eq.label..)), label.x = 3, label.y = c(12, 15)) +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")), 
           label.y = c(12, 15), 
           label.x = 6)+
  geom_point(size = 3)+ 
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = expression("Mean O"[2]~"(mg l"^-1*")"),
       y = expression("TA flux (mmol m"^-2~"h"^-1*")"), 
       color = "",
       shape = "",
       title = "Oxygen effect on TA flux") +
  scale_color_jama() +
  theme_light() +
  theme(legend.position = "bottom",
        legend.box="vertical", legend.margin=margin(),
        text = element_text(size=20))
```

*Calcifier effect on TA flux*
More TA consumption with more calcifiers. This only works if we filter out large clams from POT fresh.
```{r}
flux_data %>% filter(treatment != "AMBIENT", calcifier_tissue_g < 0.4, ta_rate_mmol_m2_h < 5) %>%
     mutate(ta_rate_mmol_m2_h = case_when(ta_rate_mmol_m2_h < -10 ~ NA,
                                        TRUE ~ ta_rate_mmol_m2_h)) %>%
 ggplot(aes(x = calcifier_tissue_g, y = ta_rate_mmol_m2_h)) +
   geom_point(size = 3, aes(color = site, shape = season)) +
   geom_smooth(method = "lm", se = FALSE, color = "gray") +
    stat_regline_equation(aes(label =  paste(..eq.label..)), label.y = 2.5, label.x = 0.04) +
   stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")), label.y = 2.1, label.x = 0.04) +
   labs(x = "Calcifier tissue (g)", 
        y = expression("TA flux (mmol m"^-2~"h"^-1*")")) +
   theme_light()
```







