---
title: "plots"
output: html_document
date: "2023-08-23"
---

```{r}
load("Data/flux_data.RData")

licor <- read_csv("Data/par.csv")

animals <- read_csv("Data/animals_epiphytes.csv") %>%
  mutate(site = paste(river, site, sep = "_"),
         id = paste(site, chamber_id, sep = "_")) %>%
  group_by(date, id, species_id) %>%
  summarize(dry_mass_g = sum(dry_mass_g, na.rm = TRUE),
            organic_g = sum(organic_g, na.rm = TRUE), 
            inorgainc_g = sum(inorganic_g, na.rm = TRUE))

sediment <- read_csv("Data/biomass.csv") %>%
  filter(ag_bg == "sediment") %>%
  mutate(id = paste(river, site, chamber_id, sep = "_")) %>%
  group_by(date, id, species) %>%
  summarize(dry_mass = sum(dry_mass_g, na.rm = TRUE),
            dry_mass_aerial = sum(biomass_g_m2, na.rm = TRUE))
```

```{r}
ggplot(flux_data, aes(x = chamber_id, y = dic_rate_mmol_m2_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  facet_grid(season~site) +
  theme_light()

ggplot(flux_data, aes(x = chamber_id, y = ta_rate_mmol_m2_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  facet_grid(season~site) +
  theme_light()

ggplot(flux_data, aes(x = chamber_id, y = o2_rate_mmol_m2_h, fill = treatment)) +
  geom_bar(stat = "identity") +
  facet_grid(season~site) +
  theme_light()
```

```{r}
ggplot(flux_data, aes(x = dic_rate_mmol_m2_h, y =ta_rate_mmol_m2_h, color = treatment)) +
  geom_point() +
  theme_light()

ggplot(flux_data, aes(x = ta_rate_mmol_g_sav_h, y =o2_rate_mmol_g_sav_h, color = treatment, shape = site)) +
  geom_point() +
  theme_light()
```

Calcs
TA changes in estuaries are related to changes in calcium ion, nitrate, ammonium, and sulfate: DTA = 2DCa2+ + DNH4+ – DNO3– – 2DSO42–
You find an oxygen increase of about 20 umol/l/h and a TA decrease of maybe 8 umol/l/h. I cannot quickly find typical O2:C ratios for SAV, so let's use a typical value of 1.5 for marine plankton. That means the DIC decline due to photosynthesis would be 13 umol/l/h. The associated DIC decline from calcification would be 8/2 = 4 umol/l/h. So I would expect a total DIC decline of 17 umol/l/h, which is about a third of what's observed.

Photosynthesis consumes ta
Calcite precipitation consumes ta
Aerobic oxidation of reduced NH4, Mn2+, Fe2+, sulfide consume ta
Aerobic respiration produces ta
Anaerobic respiration produces ta (denitrification, DNRA, sulfate reduction)
Calcite dissolution produces ta

```{r}
calcs <- flux_data %>% 
  filter(site == "YOR_BR" & season == "spring") %>%
  arrange(chamber_id) %>%
  select(chamber_id, treatment, dic_rate_mmol_m2_h, ta_rate_mmol_m2_h, o2_rate_mmol_m2_h, ph_change, max_ph) %>%
  mutate(ta_dic_ratio = ta_rate_mmol_m2_h / dic_rate_mmol_m2_h,
        dic_metab = case_when(str_detect(treatment, "L") ~ o2_rate_mmol_m2_h / -1.3,
                               str_detect(treatment, "D") ~ o2_rate_mmol_m2_h * -1),
         dic_calcif = ta_rate_mmol_m2_h / 2,
         dic_metab_calc = dic_metab + dic_calcif)
```

```{r}
ggplot(calcs, aes(x = chamber_id, y = ta_dic_ratio, fill = treatment)) +
  geom_bar(stat = "identity") +
  theme_light()
```






```{r}
#plot biomass
biomass_list <- biomass %>%
  group_split(year, season)
  
bioplot <- function(x){
  x %>%
  mutate(ag_bg = case_when(ag_bg == "ag" ~ "above ground",
                           ag_bg == "bg" ~ "below ground"),
         species = case_when(species == "cd" ~ "Ceratophyllem demersum",
                             species == "hv" ~ "Hydrilla verticillata",
                             species == "lw" ~ "Lyngbya spp.",
                             species == "ng" ~ "Najas guadeloupensis",
                             species == "red_algae" ~ "red algae",
                             species == "cladophora" ~ "green algae",
                             species == "rm" ~ "Ruppia maritima",
                             species == "zm" ~ "Zostera marina",
                             species == "unknown" ~ "mixed species roots",
                             species == "va" ~ "Vallisneria americana",
                             species == "ms" ~ "Myriophyllem spicatum",
                             species == "pper" ~ "Potamogeton perfoliatus",
                             species == "ppus" ~ "Potamogeton pusilus",
                             species == "hd" ~ "Heteranthera dubia",
                             species == "nm" ~ "Najas minor")) %>%
    filter(ag_bg != "sediment") %>%
ggplot(aes(x = chamber_id, y = biomass_g_m2, fill = species)) +
  geom_bar(stat = "identity") +
  facet_grid(ag_bg ~site) +
  labs(x = "Chamber ID", y = expression("SAV biomass (g m"^-2*")"),
       title = "Macrophyte biomass") +
  theme_light()}

map(biomass_list, bioplot)
```
